### 2024-05-29 README.Rmd
Continued writing here and there

Notice that this statement (like the definition of the work before the panels), takes for granted that 1) AIs are powerful tools, and 2) universities can be leading players in their adoption. This attitude is almost fatalistic in its enthusiasm, and perfectly consistent with the hype observed around AI this time^[The pattern of cyclical hypes around AI even has a name: [AI winters](https://en.wikipedia.org/wiki/AI_winter); the current one with chatbots in the pole position is just the most recent of several hype cycles [see @siegel_ai_2023]] [@angwin_press_2024].

@@

A possible reason why many organizations have decided to spend substantial resources considering the adoption of AIs is that, unlike most other tools, they are seen as _everything tools_. While hammers are understood to have a limited range of uses having to do with concentrating force, and word processors are mostly used to record words, the release of GenAIs to the masses was accompanied by innumerable lofty promises^[Quite a few of them made by entities and individuals that had a stake in the adoption of the technology, ranging from "AI is inevitable, and I will sell it to you" to "AI is inevitable, and for a consulting fee I will tell you how to level up with it".].

@@

@@

@@

A benefit of portraying AI as an _everything tool_ (for the proponents of the technology, at least), is that conversations around it quickly turn into a [Gish Gallop](https://en.wikipedia.org/wiki/Gish_gallop), a rhetorical device where a party makes an overwhelming number of arguments with little regard to their accuracy or strength. In the case of AI, critics are forced to respond to every single propounded use case with in-depth analysis, which usually is more time consuming than making all sorts of claims to begin with^[This communication strategy is what Trump's erstwhile and possibly future advisor Steve Bannon termed [flooding the zone with shit](https://www.vox.com/policy-and-politics/2020/1/16/20991816/impeachment-trial-trump-bannon-misinformation).].

@@

In what follows, I do not plan to confront every single possible use case of AI as doing so plays into the intent of the Gish Gallop^[I have already spent close to 50 hours working on this essay, and it is still nowhere near an exhaustive apology or indictment of AI.], which is to defeat criticism by flooding the zone. As well, I reject the dilemma posed by the purveyors of hype, and I remain convinced that there are other ways for teaching and learning to be excellent that do not require AI.

@@

Training data, much like technology, are not politically neutral, and several processes contribute to reproduce malignant biases, A way to explain this is by means of the concept of _platformed racism_ [@matamoros-fernandez_platformed_2017]. Platformed racism is the process whereby the functionalities of platforms are exploited to create and spread racist ideology (liking and sharing), often in a grass-roots fashion, but more ominously with the assist of bad-faith actors^[Online Russian disinformation has had a wide reach [@treyger_russian_2022] and furthermore is known to exploit "all kinds of separatism and ethnic, social and racial conflicts, actively supporting all dissident movements--extremist, racist, and sectarian groups" to destabilize internal U.S. politics [@us_department_of_state_gec_2024, p. 51]. Given how cagey corporations are about the data they use, it is anyone's guess how much dis- and misinformation is in the datasets used to train their AIs (the Unfathomable Data problem discussed by @bender_dangers_2021).]. The second aspect of platformed racism is the implicit endorsement of racist ideology through vague standards and [arbitrary moderation practices](https://www.bbc.com/news/technology-58058428) that exclude some views while allowing others to thrive [@myers_west_censored_2018; @hawkins_race_2023].

Search engines were already known to reinforce racism [@noble_algorithms_2018]; given the above, it is too mych to expecet that AI-augmented search engines do better^[A report by the Center for Countering Digital Hate found that Google's Bard generated wrong and/or hateful outputs 78 out of 100 tests [@center_for_countering_digital_hate_googles_2023]. In addition to being misleading to the unadvised, it is also a stupendous tool ripe for abuse by the advised with an agenda.]. In another wretched incident, Google's AI overview stated that Barack Obama was the first Muslim president of the US, a regrettable piece of misinformation that fails to surprise given the years-long right-wing disinformation campaign to otherize Obama [@tenbarge_glue_2024]. Mishaps like this, and others, are dismissed by Google as attempts to "to trip up the technology with uncommon questions" [@tenbarge_glue_2024], which only reinforces the notion that AIs only are useful when [regressing to the mean](#really-what-do-they-do). Responses like fiddling with the model (fine tune it) to provide more palatable outputs (like in the case of **FIND EXAMPLE**) are not more reassuring. Who is to say what a handful of unaccountable corporations will find palatable tomorrow, let alone in the face of powerful authoritarian regimes?

@@

>>>>

Notice that this statement (like the definition of the work before the panels), takes for granted that 1) AIs are powerful tools, and 2) universities can be leading players in their adoption. This attitude is almost fatalistic^[In the sense of "fated to"] in its enthusiasm, and perfectly consistent with the hype observed around AI this time^[The pattern of cyclical hypes around AI even has a name: [AI winters](https://en.wikipedia.org/wiki/AI_winter); the current one with chatbots in the pole position is just the most recent of several hype cycles [see @siegel_ai_2023]] [@angwin_press_2024].

@@

A key reason why many organizations have decided to spend substantial resources considering the adoption of AIs is that, unlike most other tools, they are seen as _everything tools_. While hammers are understood to have a limited range of uses having to do with concentrating force, and word processors are mostly used to record words, the release of GenAIs to the masses was accompanied by innumerable lofty promises^[Quite a few of them made by entities and individuals that had a stake in the adoption of the technology, ranging from "AI is inevitable, and I will sell it to you" to "AI is inevitable, and for a consulting fee I will tell you how to level up with it".].

@@

AI would even write code, perhaps helping itself to some self-programming (AI self-help?) as an outcome.

Figure @\ref(fig:ai-writing-code)

```{r ai-writing-code, out.width="50%", fig.cap="\\label{fig:ai-bug-fix}Figure X. Example of an AI writing code"}
knitr::include_graphics("images/AI-bug-fix.png")
```

@@

@contact_north_five_2024^[teachonline.ca by [Contact North/Contact Nord](https://teachonline.ca), is "a resource for post-secondary faculty and instructors...to find the latest information on new technology and new developments in online learning, as well as practical tools and resources to help them integrate technology in their teaching in a way that improves the learning experience for their students". Contact North/Contact Nord is a not-for-profit corporation funded by the Government of Ontario, the province that has underfunded its post-secondary education sector for decades [@usher_state_2023].] **MORE HYPE**

@@

A benefit of portraying AI as an _everything tool_ (for the proponents of the technology, at least), is that conversations around it quickly turn into a [Gish Gallop](https://en.wikipedia.org/wiki/Gish_gallop), a rhetorical device where a party makes an overwhelming number of arguments with little regard to their accuracy or strength. In the case of AI, critics are forced to respond to every single propounded use case with in-depth analysis, which usually is much more time consuming than making all sorts of claims to begin with^[This communication strategy is what Trump's erstwhile and possibly future advisor Steve Bannon termed [flooding the zone with shit](https://www.vox.com/policy-and-politics/2020/1/16/20991816/impeachment-trial-trump-bannon-misinformation). Perhaps not coincidentally, AI is now a major vector of disinformation, flooding the zone at a rate that humans cannot possibly match [@maiberg__google_2024].].

@@

In what follows, I do not plan to confront every single possible use case of AI as doing so plays into the intent of the Gish Gallop^[I have already spent close to 50 hours working on this essay, and it is still nowhere near an exhaustive apology or indictment of AI.], which is to defeat criticism by flooding the zone. As well, I reject the dilemma posed by the purveyors of hype, and I remain convinced that there are multiple other ways for teaching and learning to be excellent that do not require AI, and do not require the moral, environmental, social compromises demanded by AI[^Most discussions of AI as an everything tool acknowledge ].

@@

Training data, much like technology, are not politically neutral, and several processes contribute to reproduce malignant biases, A way to explain this is by means of the concept of _platformed racism_ [@matamoros-fernandez_platformed_2017]. Platformed racism is the process whereby the functionalities of platforms are exploited to create and spread racist ideology (liking and sharing), often in a grass-roots fashion, but more ominously with the assist of bad-faith actors^[Online Russian disinformation has had a wide reach [@treyger_russian_2022] and furthermore is known to seed distrust between ethnic communities [@svedkauskas_analysis_2020]. As the U.S. Department of State Reports, Russian intelligence is known to exploit "all kinds of separatism and ethnic, social and racial conflicts, actively supporting all dissident movements--extremist, racist, and sectarian groups" to destabilize internal U.S. politics [@us_department_of_state_gec_2024, p. 51].]. Given how cagey corporations are about the data they use to train their models, it is anyone's guess how much dis- and misinformation has been passed to their AIs (this is part of the Unfathomable Data problem discussed by @bender_dangers_2021). In this way, AIs, rather than contributing to online safety, may end up amplifying the legacy of Russian interference [@svedkauskas_analysis_2020].]. The second aspect of platformed racism is the implicit endorsement of racist ideology through vague standards and [arbitrary moderation practices](https://www.bbc.com/news/technology-58058428) that exclude some views while allowing others to thrive [@myers_west_censored_2018; @hawkins_race_2023].

Search engines were already known to reinforce racism [@noble_algorithms_2018]; given the above, it really is too much to expect that AI-augmented search engines will do better^[A report by the Center for Countering Digital Hate found that Google's Bard generated wrong and/or hateful outputs 78 out of 100 tests [@center_for_countering_digital_hate_googles_2023]. In addition to being misleading to the unadvised, AI is also a stupendous tool ripe for abuse by those who are advised _and_ have an agenda. Google's own researchers have come to the inevitable conclusion that it is easier to produce disinformation with AI, than using AI to identify misinformation [@maiberg__google_2024].]. In another wretched incident, Google's AI overview stated that Barack Obama was the first Muslim president of the US, a regrettable piece of misinformation that fails to surprise given the years-long Russian-assisted right-wing propaganda campaign to otherize Obama [@tenbarge_glue_2024]. Mishaps like this, and others, are dismissed by Google as attempts to "to trip up the technology with uncommon questions" [@tenbarge_glue_2024]. This dismissal, though, only reinforces the notion that AIs are only ever useful when [regressing to the mean](#really-what-do-they-do). Responses like fiddling with the model (fine tuning it) to provide more palatable outputs (like in the case of **FIND EXAMPLE**) are not more reassuring. Who is to say what a handful of unaccountable corporations will find palatable tomorrow, let alone in the face of powerful authoritarian regimes?

@@

What do AIs do? For the province, the main patron and effectively the boss of McMaster, it provides a rationale (an excuse even) to force the institution to find "efficiencies".
